### 2014-11-13 evaluation.tex
hfill before qed

\begin{figure*}
  \vspace{-5pt}
  \centering
  \subfigure[Bank account]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-tp-vs-lat.pdf}}
  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
  \subfigure[RUBiS bidding mix]{\label{grf:rubis}\includegraphics[width=0.24\textwidth]{graphs/Rubis.pdf}}
  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
	\caption{Quelea Performance.}
  \label{grf:LWW_perf}
\end{figure*}

Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations
in the bank account example as we increase the number of clients in a \cf{1DC}
configuration. Our client workload was generated using the YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

@@

@@

We also compared (not shown) the performance of EC LWW operations directly
against Cassandra, which uses last-writer-wins as the only convergence
semantics. While Cassandra provides no stronger-than-eventual consistency
properties, \name was within 30\%(20\%) of latency(throughput) of Cassandra
with 512 clients, supporting our thesis that programmers only have to incur
relatively low overhead for a more expressive programming model which provides
stronger provable consistency guarantees.

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
configuration against a single replica ({\sf\small NoRep}) and a strongly
replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
default RUBiS bidding mix, which has 15\% read-write interactions, which is
representative of the auction workload.  Without replication, {\sf\small NoRep}
trivially provides strong consistency. However, this deployment does not scale
beyond 1750 operations per second. Strong replication offers better throughput
at the cost of greater latency due to inter-replica coordination. The \name
deployment offers the benefit of replication, while only paying the cost of
coordination when necessary.

>>>>

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:BA-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-lat.pdf}}
  \subfigure[Throughput]{\label{grf:BA-tp}\includegraphics[width=0.32\textwidth]{graphs/BA-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-tp-vs-lat.pdf}}
	\caption{Bank account performance.}
  \label{grf:BA}
\end{figure*}

Figure~\ref{grf:BA} shows the performance of operations in bank account example
as we increase the number of clients in \cf{1DC} configuration.. Our client
workload was generated using YCSB benchmark~\cite{YCSB}. The benchmark
uniformly chose from 100,000 keys, where the operation spread was 25\%
withdraw, 25\% deposit and 50\% getBalance, which corresponds to the default
50:50 read:write mix in YCSB. We increased the number of clients from 128 to
1024, and each experiment ran for 180 seconds.

@@

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:LWW-txn-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-lat.pdf}}
  \subfigure[Throughput]{\label{grf:LWW-txn-tp}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:LWW-txn-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp-vs-lat.pdf}}
	\caption{LWW register transaction performance.}
  \label{grf:LWW-txn}
\end{figure*}

@@

We also compared (not shown) the performance of EC LWW operations directly
against Cassandra (our backing store), which uses last-writer-wins as the only
convergence semantics. While Cassandra provides no stronger-than-eventual
consistency properties, \name was within 30\%(20\%) of latency(throughput) of
Cassandra with 512 clients, illustrating that the programmers only have to pay
a minimal overhead for the expressive and stronger \name programming model.

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:rubis-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-lat.pdf}}
  \subfigure[Throughput]{\label{grf:rubis-tp}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:Rubis-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp-vs-lat.pdf}}
	\caption{Rubis bidding mix performance.}
  \label{grf:rubis}
\end{figure*}

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
configuration against a single replica (NoRep) and strongly replicated
(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
15\% read-write interactions, which is representative of the auction workload.
Without replication, NoRep trivially provides strong consistency. However, this
deployment does not scale beyond 1750 operations per second. Strong replication
offers better throughput at the cost of greater latency due to inter-replica
coordination. \name deployment offers the benefit of replication, while only
paying the cost of coordination when necessary.

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{graphs/summarization.pdf}
\caption{Impact of summarization.}
\label{grf:summarization}
\end{figure}

### 2014-11-13 implementation.tex
Added basic axioms

\begin{figure}
\begin{center}
\includegraphics[width=0.7\columnwidth]{Figures/ImplModel}
\end{center}
\caption{Implementation Model.}
\label{fig:impl_mod}
\end{figure}

@@

\subsection{Shim Layer}

@@

\subsection{Summarization}

The main challenge in realizing an efficient implementation of operation-based
replicated data types is that the state of the object defined in terms of its
set of effects grows with every effectful operation. If left unchecked,
performance slows down over time, until the shim layer memory or backing store
disk runs out of memory. Luckily, the state of the operation-based replicated
data type can often be summarized to an \emph{observably equivalent} smaller
state. For example, a last-writer-wins register with multiple updates can be
summarized to an equivalent register with only the last write. Similarly, a
bank account with a series of deposits and updates with balance $b$ is
equivalent to a bank account with a single deposit of $b$. A set with a
collection of add and remove operations is equivalent to a set with a series of
addition of live elements from the original set.

Since the semantics of summarization is specific to each data type, \name
expects a summarization function (\cf{summarize}) of type \cf{[e]->[e]} for
each RDT, where the result is smaller than the argument but is observably
equivalent. We utilize the \cf{summarize} function to summarize the object
state both in the shim layer node and the backing store, typically when the
number of effects on an object crosses a tunable threshold. Shim layer
summarization is straight-forward; a summarization thread takes the local lock
on the object, and replaces its state with the summarized state. The shim layer
node remains unavailable for that particular object during summarization
(usually a few milliseconds).

Performing summarization in the backing store is more complicated since the
whole process needs to be atomic from a client's perspective, but Cassandra
does not provide multi-row transactions. We have engineered and implemented
a scalable summarization mechanism for the backing store that permits
concurrent client operations, but nonetheless prohibits these operations
from witnessing intermediate states of the summarization process.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{Figures/ImplModel}
\end{center}
\caption{Implementation Model.}
\label{fig:impl_mod}
\end{figure}

@@

@@

We utilize the \cf{summarize} function (\S~\ref{sec:summarize}) to summarize
the object state both in the shim layer node and the backing store, typically
when the number of effects on an object crosses a tunable threshold. Shim layer
summarization is straight-forward; a summarization thread takes the local lock
on the cached object, and replaces its state with the summarized state. The
shim layer node only remains unavailable for that particular object during
summarization (usually a few milliseconds). Performing summarization in the
backing store is more complicated since the whole process needs to be atomic
from a client's perspective, but Cassandra does not provide multi-row
transactions. We have engineered and implemented a scalable summarization
mechanism for the backing store that permits concurrent client operations, but
nonetheless prohibits these operations from witnessing intermediate states of
the summarization process.

### 2014-11-13 related.tex

Program analysis has been proposed to relieve the annotation burden associated
with consistency classification.~\cite{Calm} identifies code regions that
require injection of coordination.~\cite{Sieve} automatically classifies
operations into strong and eventual consistency.~\cite{IConfluence} defines the
conditions under which an operation requires coordination. These works focus of
coarse-grained classification (eventual or strong), and do not focus on
transaction consistency. Indeed, we do see these approaches as a next step in
\name to automatically infer the contracts.

>>>>

Several conditions have been proposed to judge whether an operation on
a replicated data object needs coordination or not. ~\cite{Calm}
defines \emph{logical monotonicity} as a sufficient condition for
coordination freedom, and proposes a consistency analysis that marks
code regions performing non-monotonic reasoning (eg: aggregations,
such as \cf{COUNT}) as potential coordination points.
~\cite{IConfluence} and ~\cite{Sieve} define \emph{invariant
confluence} and \emph{invariant safety}, respectively, as conditions
for safely executing an operation without coordination. ~\cite{Sieve}
also proposes a program analysis that conservatively marks certain
operations as \emph{blue} (coordination not required), while marking
the remaining as \emph{red} (coordination required). Unlike \name,
these works focus on a coarse-grained classification of consistency as
eventual or strong, and do not focus on finer-grained transaction
isolation levels. However, the analyses they propose relieve
programmers of the burden to tag operations with consistency levels.
Indeed, we do consider automatic inference of consistency contracts
from application-specific integrity constraints as the next step for
\name.
